### 2018-10-04 BatangToru/Gibbon Abundance.Rmd
Added data and exploratory analysis

---
title: "Gibbon Abundance"
author: "Cal"
date: "23/06/2017"
output: html_document
---

```{r}
Data = read.csv('GibbonSiamang.csv',header=T) # load the data
Data = subset(Data, Data$Species == 1)
x = Data$PP.Distance ; y = Data$Forward.Distance
Data$X = Data$X.1 = Data$X.2 = NULL # Remove blank rows in CSV

# We need to rename some of the columns to work with the LT2D package:
Data$stratum = Data$Region.Label ; Data$Region.Label = NULL
Data$Height = NULL # We don't use this in this analysis 
Data$Effort = NULL 
Data$transect = Data$Sample.Label ; Data$Sample.Label = NULL
Data$location = Data$Location..0.5.4.5.km. ; Data$Location..0.5.4.5.km. = NULL
```

The Gibbon effort column seems useless...? It describes constant effort, but 
clearly isn't an indication of transect line length. Is it a description of 
time spent on the transect? In which case 52 minutes seems very arbitrary.

My suspicion is that there was one very long transect, which was split into 
four. Here's why:

```{r}
trans1 = subset(Data, transect == 1 ) ; T1 = range(trans1$location)
trans2 = subset(Data, transect == 2 ) ; T2 = range(trans2$location)
trans3 = subset(Data, transect == 3 ) ; T3 = range(trans3$location)
trans4 = subset(Data, transect == 4 ) ; T4 = range(trans4$location)
T1 ; T2 ; T3 ; T4 
```

I suspect the long transect wasn't actually cut into four equal lengths. I'll 
use the average difference between the end of one transect and the start of the 
next as my best guess for how much was right truncated of the last transect. For 
some reason the transect with the highest distance has label 1, but this isn't a
big deal. I'll place the end point of transects at the halfway point of the
final detection of one transect, and the first of the transect ahead of it (
in distance, not label).

```{r}
T1.2Gap = abs(T2[2]-T1[1])
T2.3Gap = abs(T3[2]-T2[1])
T3.4Gap = abs(T4[2]-T3[1])

differences = c(0.5, T3.4Gap, T2.3Gap, T1.2Gap)
TopTruncation = mean(differences)/2

# Create a small numberline to illustrate the point:

## define plot data
xlim <- c(0,5)
ylim <- c(0,1)
px <- rev(c(T1,T2,T3,T4)) # because names are opposite of order
py <- rep(0,length(px))
lx.buf <- 5
lx <- seq(xlim[1]+lx.buf,xlim[2]-lx.buf,len=length(px))
ly <- 20

## create basic plot outline
par(xaxs='i',yaxs='i',mar=c(5,1,1,1));
plot(NA,xlim=xlim,ylim=ylim,axes=F,ann=F);
axis(1)

## plot elements
draw.points = function(px,py,pch){
  number.of.transects = length(px)/2
  points(px,py,col='red',pch=pch, xpd=F)
}

segments(px,py,lx,ly);
draw.points(px,py,pch=3)
```