### 2018-10-04 BatangToru/KnittedGibSia.Rmd

---
title: "GibSia"
author: "Cal F-J, under supervision of Prof. David L Borchers"
date: "June 2017"
output: html_document
---
```{r, echo=F, warning=F}
library(plotrix) # load the library needed to draw the arc on our graph
```
## Abstract
We perform a 2D detection function distance sampling analysis of a Gibbon and 
Siamang dataset (73 observations), using a development version of the LT2D
package. We modify the likelihood proposed by Borchers and Cox (2016), to
account for datasets for which detections with small radial distances have been
rounded down to zero in the field. We estimate $\hat{p}$ to be XX, with 95% CI().

## Examination of the dataset
```{r}
library('DEVDEV5') # load the development version of the LT2D package
Data = read.csv('GibbSia.csv',header=T) # load the x and y data
x = Data$PP.Distance ; y = Data$Forward.Distance # 73 observations
```

```{r, echo=FALSE}
# We produce some preliminary plots to see what the data look like: 
par(mfrow=c(1,3))

# we want to add a small amount of noise to data points at (0,0) to make it 
# clearer in the graph that there is a cluster here:
x_to_plot = x                                      # create vectors to 
y_to_plot = y                                      # fill with values

for (i in (1:length(x))){                          # loop through and add noise
  if (x[i]==0 & y[i]==0){                          # where it's required
    x_to_plot[i] = x[i] + runif(1,0,0.002)
    y_to_plot[i] = y[i] + runif(1,0,0.002)
  }         
}

pdlab="Perpendicular distance"
fdlab="Forward distance"
plot(x_to_plot,y_to_plot,pch="+",ylab=fdlab,xlab=pdlab)
rmin = sort(unique(Data$Radial.Distance))[2] # Calculate Rmin
draw.arc(0,0,radius=rmin,deg1=0,deg2=360,col="red",lwd=1)
legend(0.0113,0.065,"Radius Rmin",lty=1,col="red")
hist(y,breaks=seq(0,max(na.omit(y)),length=9),xlab=fdlab,main="")
hist(x,breaks=seq(0,max(na.omit(x)),length=13),xlab=pdlab,main="")
```

It would seem by inspection, that 0.05 seems a reasonable perpendicular 
truncation distance ($w=0.05$), and 0.07 seems reasonable for $ystart$; the 
furthest forward distance at which we are able to detect animals. We have a
small cluster of values at distances close to (0,0), which is likely due to the 
rounding of recorded values in the field. Points at these coordinates have had 
a small amount of noise added to them in the scatter plot, to make the cluster
more visible. We notice a clear gap in the distribution of radial distances of 
detections between 0 and 0.0132, and so we take the latter of these values
to be rmin.

Histogram of collected radial distances: 
```{r,echo=F}
R = Data$Radial.Distance
hist(R, breaks = 25, main="Radial distances of detections")
```

## Dealing with rounded distances.

The rounding down of small radial distances down to zero leads to loss of 
information; the true location of the detected animal could be at any radial
distance within $0$ to $R_{min}$, where $R_{min}$ is the value below which all
radial distances have been rounded to 0. We must reflect this uncertainty in
true location in our likelihood:

$L_R(\boldsymbol{\beta};n)=
L_U(\boldsymbol{\beta};m)\times\Bigg[ 
\frac{\int_0^{R_{min}}
{\big\{{\large S}\left(y_{min}(x),\:x\:;\:\boldsymbol{\beta}\right)
-{\large S}\left(0,\:x\:;\:\boldsymbol{\beta}\right)\big\}
\pi(x\;;\:\boldsymbol{\phi})}\: dx}
{\int_0^{R_{min}}\pi(x\:;\:\boldsymbol{\phi})\:dx}
\times\frac{1}{\int_0^w p(x\:;\:\boldsymbol{\beta})\:\pi(x\;;\:\boldsymbol{\phi}) \:dx}\Bigg]^{n-m}$

Where $L_U(\boldsymbol{\beta};m)$ is the Borchers and Cox (2016) likelihood,
evaluated at the m detections which have not been rounded down.
$y_{min}(x)$ is the forward distance at which an animal travelling from
$\infty$ to $0$ along the $y$ axis at perpendicular $x$ intersects the quarter
circle of radius $R_{min}$ and centre located at the observer's position. $S$ is
the Borchers and Cox (2016) survivor function, and all remaining parameters also
have the same definition as their presentation in Borchers and Cox (2016).

```{r, echo=F}

# We check that when using x and y values to recalculate radial distance, 
# error with repsect to the size of rmin is small - To avoid miselecting a 
# point as being rounded, when it may not have been
Error = (mean(abs(sqrt(y**2 + x**2)-Data$Radial.Distance))/rmin)*100
NumRounded = length(x[sqrt(y**2 + x**2)<rmin])

x.rounded = x[sqrt(y**2 + x**2)<rmin]
y.rounded = y[sqrt(y**2 + x**2)<rmin]
x.unrounded = x[sqrt(y**2 + x**2)>=rmin]
y.unrounded = y[sqrt(y**2 + x**2)>=rmin]
```

## Fitting some models: 
```{r}
w=0.05;ystart=0.06 # set these to use in all our fits
```

```{r, warning=F, results='hide', message=F}
# Normal bump with hazard h1: -- poor gof -- 
b=c(-6.82664896,1.09392846) ; logphi=c(-0.03503625,0.51044374)  # start values

fit.n=fityx(y,x,b=b,hr='h1',ystart=ystart,pi.x='pi.norm',logphi=logphi,0,rmin=rmin,
             w=w,hessian=TRUE,control=list(trace=0))            # fit

par(mfrow=c(1,1))
gof.LT2D(fit.n,plot=T)
plot(fit.n,xbins=20,ybins=32,smooth.fy=TRUE,addrug=TRUE) 
phatInterval(fit.n) ; phatInterval(fit.n)*w #EHSW

fit.n$p0 # p(0)
```